---
title: "Diabetes Prevalence Choropleth Map by Neighbourhood in Toronto"
author: "[Xu Fei](https://ca.linkedin.com/in/xufei1)"
date: "`r Sys.Date()`"
output: html_document
---

This is an experiment of implementing the [Wellbeing Toronto map](map.toronto.ca/wellbeing) using R.

## Diabetes data

The diabetes prevalence values are calculated by as "Percent of People With diabetes, Both sexes, Age-Adjusted". The values are taken from the [**The Wellbeing Toronto Heath Dataset**](http://opendata.toronto.ca/social.development/wellbeing/WB-Health.xlsx). The explanation on how the data were collected is [here](http://www.torontohealthprofiles.ca/a_documents/aboutTheData/3_1_AboutTheData_AdultHealthDisease_2006.pdf).

Since this file has multiple sheets and not formatted for machines, we copied the data sheet of 2008 to "./data/WB-Health-2008.csv" for later use. 

## Polygon data

First we need to download the polygon data file in the WGS84 Coordinates Reference System. The block of code below does the downloading and unzipping.
```
fileurl <- "http://opendata.toronto.ca/gcc/neighbourhoods_planning_areas_wgs84.zip"
download.file(fileurl, './data/neighbourhoods_planning_areas_wgs84.zip', method ='curl')
unzip(zipfile = "./data/neighbourhoods_planning_areas_wgs84.zip", exdir = "./")
```

The following code basically reads the shape file and diabetes data file into 2 R objects. Then we merged them into 1 file to display using Leaflet. Finally we configured the choropleth map using continuous colour gradient.

```{r, results="hide"}
library(rgdal)
```
```{r}
library(leaflet)
TOhood <- readOGR("./data", layer = "NEIGHBORHOODS_WGS84", verbose = FALSE)
TOhood.data <- read.csv("./data/WB-Health-2008.csv")
TOhood@data$AREA_S_CD <- as.integer(TOhood@data$AREA_S_CD)
TOhood2 <- merge(TOhood, TOhood.data, by.x = "AREA_S_CD", by.y = "Neighbourhood.Id", sort = FALSE)
pal <- colorNumeric(palette = "Blues",
  domain = TOhood2[['Diabetes.Prevalence']]
  )
TOhood_popup <- paste0("<strong>Neighbourhood: </strong>",
                       TOhood2@data$AREA_NAME,
                       "<br><strong>Diabetes Prevalence: </strong>",
                       TOhood2[['Diabetes.Prevalence']]
                       )
stamen_tiles <- "http://{s}.tile.stamen.com/toner-lite/{z}/{x}/{y}.png"
stamen_attribution <- 'Map tiles by <a href="http://stamen.com">Stamen Design</a>, under <a href="http://creativecommons.org/licenses/by/3.0">CC BY 3.0</a>. Data by <a href="http://openstreetmap.org">OpenStreetMap</a>, under <a href="http://www.openstreetmap.org/copyright">ODbL</a>.'

TOhoodMap <- leaflet(data = TOhood2) %>%
  addTiles(urlTemplate = stamen_tiles,
           attribution = stamen_attribution) %>%
  setView(lng=-79.38318, lat=43.65323, zoom = 11) %>%
  addPolygons(fillColor =
                ~pal(TOhood2[['Diabetes.Prevalence']]), 
                fillOpacity = 0.7, 
                color = "#BDBDC3", 
                weight = 1, 
                popup = TOhood_popup) %>%
  addLegend("bottomright", 
            pal = pal, 
            values =~Diabetes.Prevalence,
            title = "Diabetes Prevalence",
            opacity = 1
            )
```

Now if you click on each neighbourhood in the map below, you will be able to see the name of the neighbourhood and diabetes prevalence in the popup window!

```{r, echo=FALSE}
TOhoodMap  # Print the map
```

